### 2025-05-27 appendices/mpif/proof1.tex
Reverting Chapter 4 image captions, merge appendix from mpif_article

  Theorem~\ref{theorem:pif} is a straightforward extension of Theorem~4 of \citet{chen24} to PanelPOMP models. 
  Our approach is to express an arbitrary PanelPOMP model as a POMP model by ``stacking" the model into a long format.
  Then, Algorithm~\ref{alg:mpif} is equivalent to applying Algorithm~4 of \citet{chen24} to this stacked POMP model.
  After ensuring the necessary conditions are met, their Theorem~4 provides guarantees for this to converge to the MLE, proving the conclusions of Theorem~\ref{theorem:pif}.

@@

  \begin{proof} Let $\Theta_j^{(M)}$ denote the output of $M$ iterations of Algorithm~\ref{alg:mpif} without marginalization, and recall the basic definition of the joint density of a PanelPOMP model: 
  \begin{align}
&f_{\mathbf{X}_{0:N}, \mathbf{Y}_{1:N}}(\mathbf{x}_{0:N}, \mathbf{y}_{1:N}; \theta) =
\nonumber
\\
& \hspace{20mm} \prod_{u = 1}^Uf_{X_{u, 0}}(x_{u, 0};\theta) 
\prod_{n = 1}^{N_u}f_{Y_{u, n}|X_{u, n}}(y_{u, n} | x_{u, n} ;\theta)f_{X_{u, n} | X_{u, n-1}}(x_{u, n}|x_{u, n-1}; \theta). \label{eq:ppompSI}
\end{align}
  We would like to ``stack" the unit specific processes into a single process. 
  For convenience, we refactor the model to initialize at time $t_{u, 1}$ rather than $t_{u, 0}$ for each unit $u$ by considering the density of $X_{u, 1}$ after integrating out $X_{u, 0}$. 
  That is, we use the identity 
  $$
  f_{X_{u,1}}(x_{u, 1};\theta) = \int_{\mathcal{X}} f_{X_{u, 1}|X_{u, 0}}(x_1|x_0;\theta)f_{X_{u, 0}}(x_0;\theta)dx_0
  $$
  and use this to write Eq.~\ref{eq:ppompSI} as
  \begin{align}
  f_{\mathbf{X}_{1:N}, \mathbf{Y}_{1:N}}(\mathbf{x}_{1:N}, \mathbf{y}_{1:N}; \theta) = \prod_{u = 1}^U \prod_{n = 1}^{N_u}f_{Y_{u, n}|X_{u, n}}(y_{u, n} | x_{u, n} ;\theta)f_{X_{u, n} | X_{u, n-1}}(x_{u, n}|x_{u, n-1}; \theta), \label{eq:ppompSI1}
  \end{align}
  using the convention that for $n = 1$, $f_{X_{u, n} | X_{u, n-1}}(x_{u, n}|x_{u, n-1}; \theta) = f_{X_{u,1}}(x_{u, 1};\theta)$.

  Now let $T_u = \sum_{k = 1}^u (t_{k, N_k} - t_{k, 1})$, with $T_0 = 0$, and $S_u = \sum_{k = 1}^u N_k$ for $u \in 1:U$ and $S_0 = 0$. 
  We then define new latent processes $\tilde{X}$ and $\tilde{Y}$. First,
  \begin{align}
    \tilde{X}(t) &= X_u(t_{u, 1} + (t - T_u)), \, \text{for} \,\, T_{u-1} \leq t \leq T_u.
  \end{align}
  Letting $t'_n = t_{u, n-S_u}$ for $S_{u - 1} < n \leq S_u$, the latent and observable processes are equivalent to
  \begin{align*}
  \begin{split}
    \tilde{X}_n = \tilde{X}(t'_n) = X_{u, n - S_{u - 1}} \\
    \tilde{Y}_n = Y_{u, n - S_{u-1}}. 
  \end{split}
  \end{align*}
  With this new definition of $\tilde{X}$ and $\tilde{Y}$, we have a new POMP model which has joint density
  \begin{align}
&f_{\tilde{X}_{1:S_U}, \tilde{Y}_{1:S_U}}(x_{1:S_U}, y_{1:S_U}; \theta) =  \nonumber
\\
& \hspace{20mm}
f_{\tilde{X}_1}(x_1;\theta)f_{\tilde{Y}_1|\tilde{X}_1}(y_1 | x_1; \theta)\prod_{n = 2}^{S_{U}}f_{\tilde{Y}_n|\tilde{X}_n}(y_n | x_n; \theta)f_{\tilde{X}_{n}|\tilde{X}_{n - 1}}(x_n | x_{n - 1}; \theta). \label{eq:lowPOMP}
  \end{align}

  As defined, Eqs.~\ref{eq:lowPOMP} and \ref{eq:ppompSI} describe the same model. 
  We refer to Eq~\ref{eq:lowPOMP} as the \emph{long} format, which is indicative that the model has been expressed as a low-dimensional POMP model with observation times ranging from $1$ to $\sum_{u = 1}^U N_u$, rather than a collection (or product) of POMP models, each with observation times $N_u$ for $u \in \seq{1}{U}$. 
  This representation allows us to naturally extend the theoretical framework of \citet{chen24} to the PanelPOMP model.
  Specifically, it is easily verified that Assumptions~\ref{assumption:1}--\ref{assumption:2} outlined in Appendix~\ref{sec:assumptions} ensure that the density in Eq.~\ref{eq:lowPOMP} corresponds to a state-space model that satisfies the MLE conditions of \citet{chen24}. 
  In a similar fashion, Assumption~\ref{assumption:kernel} ensures that an iterated filtering algorithm with perturbation kernel that satisfies these conditions implies the conclusion of Theorem~\ref{theorem:pif}, as it is a direct application of Theorem~4 of \citet{chen24} to this model.

    \end{proof}

@@

    For instance, it can be shown following Proposition~S2 of \citet{chen24} that choosing the perturbation density $h_{u, n}(\theta|\varphi; \sigma_{u, m})$ in lines~\ref{line:startu} and \ref{line:perturbations} of Algorithm~\ref{alg:mpif} to be a multivariate normal density satisfies Assumption~\ref{assumption:kernel}, where $\varphi$ is the mean of the distribution and, for some full rank positive definite matrix $\Sigma_0$ and sequence $\sigma_{u, m} = o(m^{-1})$, $\sigma^2_{u, m}\Sigma_0$ is the covariance.

    If $\Sigma_0$ is full rank, however, then the perturbations will be applied to all parameters at each time step, including unit specific parameters for units $v \in \{1, \ldots, U\} / \{u\}$. 
    This results in an unnecessary amount of loss of information at each step.
    This motivates the default behavior of the PIF algorithm, which is that each unit gets an initial covariance $\Sigma_{u, 0}$, and only the columns corresponding to the shared and unit-specific parameters of interest are full rank. 
    Explicitly, if we write $\Sigma_{u, 0}^{(\phi, \psi_u)}$ as the full-rank submatrix corresponding the relevant shared and unit specific parameters, then the perturbation in line~\ref{line:perturbations} in Algorithm~\ref{alg:mpif} is equivalent to 
\begin{equation*}
  % \spacingset{1.5}
  \begin{cases}
    \big(\Phi_{n}, \Psi_{u, n} \big) \sim N\Big[\big(\Phi_{n-1}, \Psi_{u, n-1} \big), \, 
        \sigma^2_{u, m}\Sigma_{u, 0}^{(\phi, \psi_u)}\Big]
    \\
    \Psi_{v, n} = \Psi_{v, n-1} \text{ for all} \, v \in \seq{1}{U}, \, v \neq u.
  \end{cases}  
\end{equation*}
 This strategy is effective in reducing the loss of information that results from perturbing parameters when the data provides limited information. 
 However, it also results in the particle depletion issue in higher dimensions that is discussed in detail in Section~\ref{sec:depletion}. 
 \citet{chen24} propose alternative perturbation strategies that are intended to balance these competing interests in low dimensional models. 
 In the PanelPOMP models we are interested in this study, the proposed alternatives would still require a large number of perturbations of parameters that are not of direct interest, resulting in a large loss of information. 
 For this reason, the marginalization procedure is useful in practice as it directly addresses both issues simultaneously in PanelPOMP models. 
 Combining the novel perturbations strategies of \citet{chen24} with a marginalization step could also result in an improved algorithm in some cases, but that is outside of the scope of this article.

@@

One reason that Theorem~\ref{theorem:pif} cannot be used directly to infer the practicality of panel iterated filtering algorithms is because the behavior of $C_M$ as $M \rightarrow \infty$ is unknown.
Previous works on high-dimensional particle filtering---without adding perturbations or performing data cloning---suggest that the sequence $C_M$ might scale exponentially with the number of units \citep{snyder08,bengtsson08}.
While this problem has partially been avoided by writing the PanelPOMP in a long format, reducing the size of both the latent and observed spaces at each time point, the parameter space for $\Theta$ is still large.
Specifically, the total dimension of $\Theta$ is $d_\phi + Ud_u$, where $d_\phi$ and $d_u$ are the number of shared and unit-specific parameters, respectively.

>>>>

  Theorem~\ref{theorem:pif} is a straightforward extension of Theorem~4 of \citet{chen24} to PanelPOMP models. 
  Our approach is to express an arbitrary PanelPOMP model as a POMP model using a long format, following the nomenclature from the R \code{tidyverse} \citep{wickham19}.
  Then, Algorithm~\ref{alg:mpif} is equivalent to applying Algorithm~4 of \citet{chen24} to this long POMP model.
  After ensuring the necessary conditions are met, their Theorem~4 provides guarantees for this to converge to the MLE, proving the conclusions of Theorem~\ref{theorem:pif}.

@@

  \begin{proof} 
  % Let $\Theta_j^{(M)}$ denote the output of $M$ iterations of Algorithm~\ref{alg:mpif} without marginalization, and 
  Recall the basic definition of the joint density of a PanelPOMP model: 
  \begin{align}
&f_{\bm{X}_{0:N}, \bm{Y}_{1:N}}(\bm{x}_{0:N}, \bm{y}_{1:N}; \, \theta) =
\nonumber
\\
& \hspace{20mm} \prod_{u = 1}^Uf_{X_{u, 0}}(x_{u, 0};\, \theta) 
\prod_{n = 1}^{N_u}f_{Y_{u, n}|X_{u, n}}(y_{u, n} | x_{u, n} ;\, \theta)f_{X_{u, n} | X_{u, n-1}}(x_{u, n}|x_{u, n-1}; \, \theta). \label{eq:ppompSI}
\end{align}
  We would like to pivot the unit specific processes into a single long process, avoiding the need to loop over the unit $u$.
  We define $S_{\tilde{u}} = \sum_{k = 1}^{\tilde{u}}(N_k + 1)$ to be the sum total number of time-steps (observations + initialization) for units $1$ up to unit $\tilde{u}$, defining $S_0 = 0$.
  We use the sequence $\nstack \in \mathbb{N}$ is used to map time points and states indexed with $(u, n) \in \seq{1}{U}\times\seq{0}{N_u}$ to a new model with only a single index $\nstack \in \seq{1}{S_U}$, defined by the equation $\nstack = S_{\unit-1} + n + 1$
  % For convenience, we refactor the model to initialize at time $t_{u, 1}$ rather than $t_{u, 0}$ for each unit $u$ by considering the density of $X_{u, 1}$ after integrating out $X_{u, 0}$. 
  % That is, we use the identity 
  % $$
  % f_{X_{u,1}}(x_{u, 1};\, \theta) = \int_{\mathcal{X}} f_{X_{u, 1}|X_{u, 0}}(x_1|x_0;\, \theta)f_{X_{u, 0}}(x_0;\, \theta)dx_0
  % $$
  % and use this to write Eq.~\ref{eq:ppompSI} as
  % \begin{align}
  % f_{\bm{X}_{1:N}, \bm{Y}_{1:N}}(\bm{x}_{1:N}, \bm{y}_{1:N}; \, \theta) = \prod_{u = 1}^U \prod_{n = 1}^{N_u}f_{Y_{u, n}|X_{u, n}}(y_{u, n} | x_{u, n} ;\, \theta)f_{X_{u, n} | X_{u, n-1}}(x_{u, n}|x_{u, n-1}; \, \theta), \label{eq:ppompSI1}
  % \end{align}
  % using the convention that for $n = 1$, $f_{X_{u, n} | X_{u, n-1}}(x_{u, n}|x_{u, n-1}; \, \theta) = f_{X_{u,1}}(x_{u, 1};\, \theta)$.

  Now let $T_{\tilde{u}} = \sum_{k = 1}^{\tilde{u}} (t_{k, N_k} - t_{k, 0})$ for $\tilde{u} \in \seq{1}{U}$, with $T_0 = 0$.
  We now define new latent processes $\tilde{X}$ and $\tilde{Y}$. As the original model allows for continuous time latent process, we write 
  First,
  \begin{align}
    \tilde{X}(t) &= X_u(t_{u, 0} + (t - T_u)), \, \text{for} \,\, T_{u-1} \leq t \leq T_u.
  \end{align}
  Letting $\tau_{\nstack} = t_{u, n}$ using the mapping $\nstack = S_{\unit-1} + n + 1$, the latent and observable processes are equivalent to
  \begin{align*}
  \begin{split}
    \tilde{X}_{\nstack} = \tilde{X}(\tau_{\nstack}) = X_{u, n} \\
    \tilde{Y}_{\nstack} = Y_{u, n}. 
  \end{split}
  \end{align*}
  With this new definition of $\tilde{X}$ and $\tilde{Y}$, we have a new POMP model which has joint density
  \begin{align}
&f_{\tilde{X}_{1:S_U}, \tilde{Y}_{1:S_U}}(x_{1:S_U}, y_{1:S_U}; \, \theta) =  \nonumber
\\
& \hspace{20mm}
f_{\tilde{X}_0}(x_0;\, \theta)\prod_{\nstack = 1}^{S_{U}}f_{\tilde{Y}_{\nstack}|\tilde{X}_\nstack}(y_{\nstack} | x_{\nstack}; \, \theta)f_{\tilde{X}_{\nstack}|\tilde{X}_{\nstack - 1}}(x_{\nstack} | x_{\nstack - 1}; \, \theta), \label{eq:lowPOMP}
  \end{align}
  Using the convention that $n' \mapsto (u,0)$ for any $u$, then $f_{\tilde{Y}_{\nstack}|\tilde{X}_\nstack}(y_{\nstack} | x_{\nstack}; \, \theta) = 1$ and \linebreak $f_{\tilde{X}_{\nstack}|\tilde{X}_{\nstack - 1}}(x_{\nstack} | x_{\nstack - 1}; \, \theta) = f_{X_{u, 0}}(x_{u, 0}; \, \theta)$.

  As defined, Eqs.~\ref{eq:lowPOMP} and \ref{eq:ppompSI} describe the same model, but \ref{eq:lowPOMP} has been written to match the state space model of \citet{chen24}, after adjusting for the choice of initializing at $\tilde{X}_1$ rather than $\tilde{X}_0$. 
  We refer to Eq~\ref{eq:lowPOMP} as the \emph{long} format, which is indicative that the model has been expressed as a low-dimensional POMP model with observation times ranging from $1$ to $S_U$, rather than a collection (or product) of POMP models, each with observation times $N_u+1$ for $u \in \seq{1}{U}$.

  This representation allows us to naturally extend the theoretical framework of \citet{chen24} to PanelPOMP models.
  Specifically, Assumption~\ref{assumption:regular} imply that the set $\Theta$ is a regular compact set \citep[See Definition~1 of][]{chen24}; Assumptions~\ref{assumption:mle1}--\ref{assumption:mle2} ensure that the density in Eq.~\ref{eq:lowPOMP} corresponds to a state-space model that satisfies the MLE conditions of \citet{chen24}. 
  Finally, Assumptions~\ref{assumption:kernel1}--\ref{assumption:kernel4} are applied to the cloned version of Model~\ref{eq:lowPOMP}. Specifically, the process is cloned such that we have a new state space model $\hat{Y}_{\nclone} = Y_{u, n}$, $\hat{X}_{\nclone} = X_{u, n}$, using the mapping $\nclone = (\nmif-1)S_\Unit + S_{\unit-1}+n+1$.
  Assumptions~\ref{assumption:kernel1}--\ref{assumption:kernel4} therefore imply assumptions C1-C3 and C4' of \citet{chen24} for the perturbation kernels of the self organized state space model defined via $\big(\hat{Y}_{\nclone}, \hat{X}_{\nclone}\big)$.
  Together, the conditions stated in Appendix~\ref{sec:assumptions} allow for a direct application of Theorem~4 of \citet{chen24} to this model.
  % See S6.2.3 and S6.2.4 of Chen et al (2024). Here, all numbers 
  % are referencing the Chen et al (2024) paper unless otherwise stated. To use 
  % Theorem~4, which gives convergence of the particle representation, we need 
  % Propositions S10-S11, and Theorems A.1 / A.2: 
  % 
  %    - Theorems A.1 and A.2 give same results, with different conditions. 
  %      Specifically, A.2 assumes conditions C1-C3 + C4'. In our version of 
  %      the assumptions, we only give C4' (which is our assumption (C4)), so 
  %      this only gives us A.2, which is all that is needed. To further satisfy 
  %      A.2, we need: 
  %       - We need assumptions A1-A2 to hold. Assumption MLE implies A1 (by 
  %         proposition S10). We satisfy assumption MLE with our assumptions 
  %         B1-B2. 
  %       - They never say the explicitly, but A2 is satisfied by the MLE 
  %         condition. Specifically, G_{s, \theta}(x', x) = f(y|x; \theta) 
  %         (from Proposition S9), and Assumption MLE says that 
  %         sup f(y|x; \theta) is finite.
  %    - Proposition S10 is satisfied by Assumption (MLE), and the definition 
  %      of the model (Proposition S9). 
  %    - Proposition S11 is readily satisfied by the model and perturbation 
  %      definitions (Proposition S9). 
  % 
  % Our additional assumption (A1) we haven't mentioned yet is just needed to 
  % ensure the regular compact set (Definition 1 of Chen et al (2024)).
  % 
  % Finally, the assumptions C1-C3 + C4' are satisfied for Gaussian 
  % perturbations via Section S3, prop S2 (also see comment about pomp package, 
  % which gives the PIF defaults, in Section 3.3). 
    \end{proof}

@@

    For instance, it can be shown following Proposition~S2 of \citet{chen24} that choosing the perturbation density $h_{u, n}(\theta|\varphi; \sigma_{u, m})$ in lines~\ref{line:startu} and \ref{line:perturbations} of Algorithm~\ref{alg:mpif} to be a multivariate normal density satisfies Assumptions~\ref{assumption:kernel1}--\ref{assumption:kernel4}, where $\varphi$ is the mean of the distribution and, for some full rank positive definite matrix $\Sigma_0$ and sequence $\sigma_{u, m} = o(m^{-1})$, $\sigma^2_{u, m}\Sigma_0$ is the covariance.

    If $\Sigma_0$ is full rank, however, then the perturbations will be applied to all parameters at each time step, including unit specific parameters for units $v \in \{1, \ldots, U\} / \{u\}$. 
    This results in an unnecessary amount of loss of information at each step.
    This motivates the default behavior of the PIF algorithm, which is that each unit gets an initial covariance $\Sigma_{u, 0}$, and only the columns corresponding to the shared and unit-specific parameters of interest are full rank. 
    Explicitly, if we write $\Sigma_{u, 0}^{(\phi, \psi_u)}$ as the full-rank submatrix corresponding the relevant shared and unit specific parameters, then the perturbation in line~\ref{line:perturbations} in Algorithm~\ref{alg:mpif} is equivalent to 
\begin{equation*}
  \begin{cases}
    \big(\Phi_{n}, \Psi_{u, n} \big) \sim N\Big[\big(\Phi_{n-1}, \Psi_{u, n-1} \big), \, 
        \sigma^2_{u, m}\Sigma_{u, 0}^{(\phi, \psi_u)}\Big]
    \\
    \Psi_{v, n} = \Psi_{v, n-1} \text{ for all} \, v \in \seq{1}{U}, \, v \neq u.
  \end{cases}  
\end{equation*}
 This strategy is effective in reducing the loss of information that results from perturbing parameters when the data provides limited information. 
 However, it also results in the particle depletion issue in higher dimensions that is discussed in detail in Section~\ref{sec:depletion}. 
 \citet{chen24} propose alternative perturbation strategies that are intended to balance these competing interests in low dimensional models. 
 In the PanelPOMP models we are interested in this study, the proposed alternatives would still require a large number of perturbations of parameters that are not of direct interest, resulting in a large loss of information. 
 For this reason, the marginalization procedure is useful in practice as it directly addresses both issues simultaneously in PanelPOMP models. 
 Combining the novel perturbations strategies of \citet{chen24} with a marginalization step could also result in an improved algorithm in some cases, but that is outside of the scope of this article.

@@

One reason that Theorem~\ref{theorem:pif} cannot be used directly to infer the practicality of panel iterated filtering algorithms is because the behavior of $C_M$ as $M \rightarrow \infty$ is unknown.
Previous works on high-dimensional particle filtering---without adding perturbations or performing data cloning---suggest that the sequence $C_M$ might scale exponentially with the number of units \citep{snyder08,bengtsson08}.
While this problem has partially been avoided by writing the PanelPOMP in a long format, reducing the size of both the latent and observed spaces at each time point, the parameter space for $\Theta$ is still large.
Specifically, the total dimension of $\Theta$ is $\nshared + U\nspecific$, where $\nshared$ and $\nspecific$ are the number of shared and unit-specific parameters, respectively.

### 2025-05-27 chapters/mpif/ms.Rnw
Reverting Chapter 4 image captions, merge appendix from mpif_article

ggplot(df_all_mu) +
  geom_point(
    aes(x = mu1, y = mu2, col = as.factor(step)),
    pch = 18, size = 1
  ) +
  geom_path(
    data = df_all_Sigma,
    aes(x = X1, y = X2, col = as.factor(step))
  ) +
  geom_path(
    data = like_elipse,
    aes(x = X1, y = X2), linetype = 'dashed'
  ) +
  geom_point(x = 0, y = 0, col = 'red', pch = 'x') +
  theme_bw() +
  scale_x_continuous(limits = c(-3.12, 13.5)) +
  scale_y_continuous(limits = c(-3.12, 13.5)) +
  theme(aspect.ratio = 1) +
  labs(color = 'Step') +
  scale_color_manual(
    values = c(
      "#800026",
      "#bd0026",
      "#e31a1c",
      "#fc4e2a",
      "#fd8d3c",
      "#feb24c"
    )
  ) +
  labs(x = TeX('$\\phi$'), y = TeX('$\\psi$')) +
  theme(legend.position = 'right') +
  facet_wrap(~group, labeller = labeller(group = method_names)) + 
  theme(plot.margin = unit(c(0, 0, 0, 0), 'cm'))
@
\caption{\label{fig:DC}Data Cloning and Marginalized Data Cloning for two parameter model with Gaussian likelihoods and priors. The ellipses show the region of the parameter distribution that contains $95\%$ of the probability mass of the distribution. The black dashed line shows this region for the likelihood surface, and the red ``x" marks the MLE.}
\end{figure}

@@

cowplot::plot_grid(depletion_1, depletion_2, nrow = 2, labels = "AUTO", rel_heights = c(0.425, 0.575))
@
\caption{\label{fig:depletion}Updating parameter distributions with a single $u = 1$ iteration of both versions of Algorithm~\ref{alg:mpif}. (A) The total number of unique particles representing each parameter. The dashed horizontal line shows that MPIF maintains the number of particles for $\Psi_2$ over time. (B) Parameter particle swarm of a single update with and without marginalization compared to the true posterior distribution.}
\end{figure}

@@

cowplot::plot_grid(gg_smallU, gg_bigU, nrow = 2, rel_heights = c(0.4, 0.6))
@
\caption{\label{fig:gomp}Comparison of the MPIF and PIF algorithms for fitting the stochastic Gompertz population model. 
The solid horizontal line shows the true maximum likelihood, determined via the Kalman filter and a numeric optimizer, an intractable approach for high-dimensional parameter spaces $(U > 100)$; in these cases, the dashed line indicates the likelihood at the data-generating parameters. 
Each algorithm used 50 unique starting points.
Vertical bars span the tenth percentile to the maximum likelihood values; lowest values were excluded due to poor likelihoods from the PIF algorithm.}
\end{figure}

@@

print(gg)
@
\caption{\label{fig:measles1} Log-likelihoods yielded by the PIF and MPIF algorithms for the mechanistic measles model to the UK data. 
Rows correspond to a different number of particles $J$ used in Algorithm~\ref{alg:mpif}. 
The log-likelihood is evaluated at iterations 20, 56, 92, 128, 164, and 200, marked by vertical grey lines.}
\end{figure}

>>>>

ggplot(df_all_mu) +
  geom_point(
    aes(x = mu1, y = mu2, col = as.factor(step)),
    pch = 18, size = 1
  ) +
  geom_path(
    data = df_all_Sigma,
    aes(x = X1, y = X2, col = as.factor(step))
  ) +
  geom_path(
    data = like_elipse,
    aes(x = X1, y = X2), linetype = 'dashed'
  ) +
  geom_point(x = 0, y = 0, col = 'red', pch = 'x') +
  theme_bw() +
  scale_x_continuous(limits = c(-3.12, 13.5)) +
  scale_y_continuous(limits = c(-3.12, 13.5)) +
  theme(aspect.ratio = 1) +
  labs(color = 'Step') +
  scale_color_manual(
    values = c(
      "#800026",
      "#bd0026",
      "#e31a1c",
      "#fc4e2a",
      "#fd8d3c",
      "#feb24c"
    )
  ) +
  labs(x = TeX('$\\phi$'), y = TeX('$\\psi$')) +
  theme(legend.position = 'right') +
  facet_wrap(~group, labeller = labeller(group = method_names)) + 
  theme(plot.margin = unit(c(0, 0, 0, 0), 'cm'))
@
\caption[Data cloning compared to marginalized data cloning for Gaussian densities.]{\label{fig:DC}Data cloning compared to marginalized data cloning for Gaussian densities. The ellipses show the region of the parameter distribution that contains $95\%$ of the probability mass of the distribution. The black dashed line shows this region for the likelihood surface, and the red ``x" marks the MLE.}
\end{figure}

@@

cowplot::plot_grid(depletion_1, depletion_2, nrow = 2, labels = "AUTO", rel_heights = c(0.425, 0.575))
@
\caption[Updating parameter distributions with a single $u = 1$ iteration of both versions of Algorithm~\ref{alg:mpif}.]{\label{fig:depletion}Updating parameter distributions with a single $u = 1$ iteration of both versions of Algorithm~\ref{alg:mpif}. (A) The total number of unique particles representing each parameter. The dashed horizontal line shows that MPIF maintains the number of particles for $\Psi_2$ over time. (B) Parameter particle swarm of a single update with and without marginalization compared to the true posterior distribution.}
\end{figure}

@@

cowplot::plot_grid(gg_smallU, gg_bigU, nrow = 2, rel_heights = c(0.4, 0.6))
@
\caption[Comparing the MPIF and PIF algorithms for fitting the stochastic Gompertz model.]{\label{fig:gomp}Comparing the MPIF and PIF algorithms for fitting the stochastic Gompertz model.
The solid horizontal line shows the true maximum likelihood, determined via the Kalman filter and a numeric optimizer, an intractable approach for high-dimensional parameter spaces $(U > 100)$; in these cases, the dashed line indicates the likelihood at the data-generating parameters.
Each algorithm used 50 unique starting points.
Vertical bars span the tenth percentile to the maximum likelihood values; lowest values were excluded due to poor likelihoods from the PIF algorithm.}
\end{figure}

@@

print(gg)
@
\caption[Comparing the MPIF and PIF algorithms using a mechanistic model for measles.]{\label{fig:measles1} 
Comparing the MPIF and PIF algorithms using a mechanistic model for measles.
Rows correspond to a different number of particles $J$ used in Algorithm~\ref{alg:mpif}. 
The log-likelihood is evaluated at iterations 20, 56, 92, 128, 164, and 200, marked by vertical grey lines.}
\end{figure}
